The Nuclear Forensics Problem and Statistical Methods: \\ Evaluating Machine Learning for Spent Fuel Prediction

### 2017-05-18 paper.tex
adding some experimental design notes

\documentclass{anstrans}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{%
The Nuclear Forensics Problem and Statistical Methods: \\
Evaluating the Nuclear Fuel Cycle with Machine Learning
}
\author{Arrielle C. Opotowsky, Paul P.H. Wilson}

@@

Nuclear Forensics Workflow Although this focuses on Some materials of special
interest: SNF and reprocessed MOX.  For SNF, measure material, use isotope
content and/or isotope ratios to determine things like reactor type, fuel type
and enrichment at beginning of irradiation, cooling time, burnup. For
reprocessed MOX, ??.

After the material characteristics are measured, they are matched in a
forensics database that includes some or all this information for
pre-existing/pre-measured SNF/MOX. These databases are kept by individual
countries, and a given database will have widely varying uncertainty depending
where the material was measured as well as missing data in some fields.
Therefore, matching can be difficult.

@@

Add some real ML background here
\begin{subequations} \label{eqs:fullTransport}
\begin{multline} \label{eq:fullTransportVol}
  \vec{\Omega}\vd \grad \psi(\vec{x}, \vec{\Omega})
  + \sigma(\vec{x}) \psi (\vec{x}, \vec{\Omega})
\\ =
  \frac{\sigma_s(\vec{x})}{4\pi} \int_{4\pi} \psi(\vec{x},\vec{\Omega}')
  \ud\Omega' + \frac{q(\vec{x})}{4\pi}
  \equiv \frac{1}{4\pi} Q(\vec{x}) \,,
\end{multline}
inside $\vec{x} \in V$, $\vec{\Omega} \in 4\pi$, with an incident boundary
condition
\begin{equation} \label{eq:fullTransportBndy}
  \psi(\vec{x}, \vec{\Omega}) = \psi^b(\vec{x}, \vec{\Omega}) \,,
 \quad \vec{x} \in \partial V, \ \vec{\Omega} \vd \vec{n} < 0\,.
\end{equation}
\end{subequations}

@@

Talk about Dayman paper and how I'm extending that, to include neural nets
first and to eval mucking up the training data second.

@@

Preprocessing, kNN, ridge, neural net - these are all discussed above, but can
talk deets + parameters here

@@

Classification Training Error: Predetermined test set for training error,
perhaps can make a learning curve (maybe not until there is a variable
train/test set for cross validation)

Classification and Regression Error Confidence: Confidence intervals on error
(helps understand true error versus just the sample error), Test set must be >
30 instances, Can easily calculate N\% confidence interval.

Generalizability should be evaluated (learning more about this with cs760
project)

@@

Options for comparison of algorithms: Comparing classification of 2 classes on
same ROC plot with multiple ML systems, Scatter plots, Pairwise t-tests

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results and Analysis}
The results were interesting, so interesting in fact that we have decided to
present them here.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Subsection Goes Here}
The user must manually capitalize initial letters of a subsection heading.

For those who like equations in their papers, \LaTeX\ is a good choice. Here is
an equation for the Marshak diffusion boundary condition:
\begin{equation} \label{eq:marshak}
  4 J^- = \phi + 2 D \vec{n} \vd \grad \phi \,.
\end{equation}
If we so choose, we can effortlessly reference the equation later.

Another paragraph starts with Eq.~\eqref{eq:marshak} and sets $J^-$ to zero, a
vacuum boundary condition:
\begin{equation*}
  0 = \phi + \frac{2}{3} \frac{1}{\sigma} \vec{n} \vd \grad \phi \,.
\end{equation*}
The extrapolation distance is $2/3$. A more detailed asymptotic analysis yields
an extrapolation distance of about $0.71045$.

Figure~\ref{fig:voltage} shows how a plot might conceivably look in your
document. Always place figures after they are referenced so as not to throw
off the reader. You can use symbols and different line styles to help
differentiate your results, especially if they are printed in black and white.
Note how Fig.~\ref{fig:voltage} uses dashed lines \verb|--| for the exact
solution, solid lines \verb|-| for the new method's solutions, and dotted lines
\verb|:| for existing inaccurate methods.
\begin{figure}[ht] % replace 't' with 'b' to force it to be on the bottom
  \centering
  \includegraphics{example_figure}
  \caption{Captions are flush with the left.}
  \label{fig:voltage}
\end{figure}

Later on, we can include a table, even one that spans two columns such as
Table~\ref{tab:widetable}.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table*}[htb]
  \centering
\begin{tabular}{llllllllll}\toprule
      & $\phi_T(0)$      & $\phi_T(10)$      & $\phi_T(20)$      &
      $\phi_D(0)$      & $\phi_D(10)$      & $\phi_D(20)$      & $\rho$      &
      $\varepsilon$      & $N_\text{it}$
\\ \midrule
$c=0.999$  & 0.9038 & 20.63 & 31.24 & 0.9087 & 20.63 & 31.23 & 0.2192 & $10^{-7}$ & 15
\\
$c=0.990$  & 0.3675 & 13.04 & 24.7 & 0.3696 & 13.04 & 24.69 & 0.2184 & $10^{-7}$ & 15
\\
$c=0.900$  & 0.009909 & 4.776 & 17.64 & 0.009984 & 4.786 & 17.63 & 0.2118 & $10^{-7}$ & 14
\\
$c=0.500$  & $6.069\times 10^{-5}$ & 2.212 & 15.53 & 6.213$\times 10^{-5}$ & 2.239 & 15.53 & 0.2068 & $10^{-7}$ & 13
\\
\bottomrule
\end{tabular}
  \caption{This is an example of a really wide table which might not normally
  fit in the document.}
  \label{tab:widetable}
\end{table*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Notice how the table reference uses a Roman numeral
for its numbering scheme, whereas the figure reference uses an Arabic numeral.
For one-column tables, use the \verb|table| environment; two-column tables use
\verb|table*|. The same applies to figures.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Another Subsection}
Excessive sectioning in a three-page document is discouraged, but here are more
subsections to demonstrate compliance with the ANS formatting guidelines.

\subsubsection{Third-level Heading}
This subsubsection shows compliance with the ANS-specified standard. This level
of heading should be used rarely.

\subsubsection{Another Such Heading}
And, if you really think you need a third-level heading, you should make sure
that your subsection needs at least two of them.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}

The included ANS style file and this clear example file are a panacea for
the hours of headache that invariably results from formatting a document in
Microsoft Word.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\appendix
\section{Appendix}

Numbering in the appendix is different:
\begin{equation} \label{eq:appendix}
  2 + 2 = 5\,.
\end{equation}
and another equation:
\begin{equation} \label{eq:appendix2}
  a + b = c\,.
\end{equation}

>>>>

\documentclass{anstrans}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{%
The Nuclear Forensics Problem and Statistical Methods: \\
Evaluating Machine Learning for Spent Fuel Prediction
}
\author{Arrielle C. Opotowsky, Paul P.H. Wilson}

@@

Workflow for determining SNF quantities of interest: measure material, use
isotope content and/or isotope ratios to determine things like reactor type,
fuel type and enrichment at beginning of irradiation, cooling time, burnup.

After the material characteristics are measured, they are matched in a
forensics database that includes some or all this information for
pre-existing/pre-measured SNF. These databases are kept by individual
countries, and a given database will have widely varying uncertainty depending
where the material was measured as well as missing data in some fields.
Therefore, matching can be difficult.

@@

Add some real ML background here

@@

This work begins by replicating the training and test sets used in ref (cite
Dayman). The study does evaluate the imapact of randomly introduced error of
varying amounts on the ability of the algorithms to correctly predict the
burnup and reactor type. It also varies the amount of nuclides given as
training to the algorithms (a full list, a short list determined by PCA, and
fission products only). However, machine learning algorithms are heavily
dependent on the inputs and parameters given to them, such as training set
sizes, learning rates, etc. Additionally, it does not consider more advanced
aglorithms. This paper adds in the comparison of the simpler algorithms to
neural nets. It also presents learning curves for each algorithm to investigate
the effect of the training set size on prediction error. For the classification
of reactor type, ROC curves are also presented to evaluate prediction and
generalization strength. For the regression of burnup, ??? pairwise t test?.
Thus, I will first investigate these algorithms with no error introduced, but
will extend the validation from a predetermined test set to be randomly
partitioned cross validation.  And the next study will introduce error by
limiting the nuclides to only those that can be measured with a gamma
spectrometer (future work).

@@

NN classifier+regressor, ridge classification + regression, neural net
classification + regression- these are all discussed above, but can talk deets
+ parameters here

@@

Using predetermined test set:

1. Learning curves: For a given (randomly chosen) training set size for each
algorithm, run several trainings and predictions. This helps determine if we
are over or under training and each algorithm's robustness to over/under
fitting.

2. Classification Training Error: Precision/recall values / ROC curves for the
3 classes or maybe comparing algorithms to each other for each class

3. Regression Training Error: Confidence intervals on predictions to understand
true error versus sample error Test set must be > 30 instances, Can easily
calculate N\% confidence interval.

Using cross-validation instead:

1. redo learning curve study without having to run several trainings and
predictions for each set (bc that's already done!)

2. redo precision/recall study

3. redo confidence interval study

@@

Maybe can delete this if I already discuss it in the above section. Options for
comparison of algorithms: Comparing classification of 2 classes on same ROC
plot with multiple ML systems, Scatter plots, Pairwise t-tests.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results and Analysis}

So far, should include, with all 4 algorithms shown in each plot:
1. Learning curves  + CV curve
2. ROC plots + CV plot
3. Confidence interval tables + CV table

For the above categories in the validation section, can show plots directly
next to each other (for 1, 2, 3) to hopefully show that cross validation
provides better generalizability.  Future work would be to investigate the
number of cross validation folds.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsection{Subsection Goes Here}
%The user must manually capitalize initial letters of a subsection heading.
%
%For those who like equations in their papers, \LaTeX\ is a good choice. Here is
%an equation for the Marshak diffusion boundary condition:
%\begin{equation} \label{eq:marshak}
%  4 J^- = \phi + 2 D \vec{n} \vd \grad \phi \,.
%\end{equation}
%If we so choose, we can effortlessly reference the equation later.
%
%Another paragraph starts with Eq.~\eqref{eq:marshak} and sets $J^-$ to zero, a
%vacuum boundary condition:
%\begin{equation*}
%  0 = \phi + \frac{2}{3} \frac{1}{\sigma} \vec{n} \vd \grad \phi \,.
%\end{equation*}
%The extrapolation distance is $2/3$. A more detailed asymptotic analysis yields
%an extrapolation distance of about $0.71045$.
%
%Figure~\ref{fig:voltage} shows how a plot might conceivably look in your
%document. Always place figures after they are referenced so as not to throw
%off the reader. You can use symbols and different line styles to help
%differentiate your results, especially if they are printed in black and white.
%Note how Fig.~\ref{fig:voltage} uses dashed lines \verb|--| for the exact
%solution, solid lines \verb|-| for the new method's solutions, and dotted lines
%\verb|:| for existing inaccurate methods.
%\begin{figure}[ht] % replace 't' with 'b' to force it to be on the bottom
%  \centering
%  \includegraphics{example_figure}
%  \caption{Captions are flush with the left.}
%  \label{fig:voltage}
%\end{figure}
%
%Later on, we can include a table, even one that spans two columns such as
%Table~\ref{tab:widetable}.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\begin{table*}[htb]
%  \centering
%\begin{tabular}{llllllllll}\toprule
%      & $\phi_T(0)$      & $\phi_T(10)$      & $\phi_T(20)$      &
%      $\phi_D(0)$      & $\phi_D(10)$      & $\phi_D(20)$      & $\rho$      &
%      $\varepsilon$      & $N_\text{it}$
%\\ \midrule
%$c=0.999$  & 0.9038 & 20.63 & 31.24 & 0.9087 & 20.63 & 31.23 & 0.2192 & $10^{-7}$ & 15
%\\
%$c=0.990$  & 0.3675 & 13.04 & 24.7 & 0.3696 & 13.04 & 24.69 & 0.2184 & $10^{-7}$ & 15
%\\
%$c=0.900$  & 0.009909 & 4.776 & 17.64 & 0.009984 & 4.786 & 17.63 & 0.2118 & $10^{-7}$ & 14
%\\
%$c=0.500$  & $6.069\times 10^{-5}$ & 2.212 & 15.53 & 6.213$\times 10^{-5}$ & 2.239 & 15.53 & 0.2068 & $10^{-7}$ & 13
%\\
%\bottomrule
%\end{tabular}
%  \caption{This is an example of a really wide table which might not normally
%  fit in the document.}
%  \label{tab:widetable}
%\end{table*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Notice how the table reference uses a Roman numeral
%for its numbering scheme, whereas the figure reference uses an Arabic numeral.
%For one-column tables, use the \verb|table| environment; two-column tables use
%\verb|table*|. The same applies to figures.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsection{Another Subsection}
%Excessive sectioning in a three-page document is discouraged, but here are more
%subsections to demonstrate compliance with the ANS formatting guidelines.
%
%\subsubsection{Third-level Heading}
%This subsubsection shows compliance with the ANS-specified standard. This level
%of heading should be used rarely.
%
%\subsubsection{Another Such Heading}
%And, if you really think you need a third-level heading, you should make sure
%that your subsection needs at least two of them.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\section{Conclusions}
%
%The included ANS style file and this clear example file are a panacea for
%the hours of headache that invariably results from formatting a document in
%Microsoft Word.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\appendix
%\section{Appendix}
%
%Numbering in the appendix is different:
%\begin{equation} \label{eq:appendix}
%  2 + 2 = 5\,.
%\end{equation}
%and another equation:
%\begin{equation} \label{eq:appendix2}
%  a + b = c\,.
%\end{equation}
